---
title: "Hypoxic Gene Expression"
author: "Jannik Hannemann"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  ioslides_presentation:
    incremental: true
    widescreen: true 
    logo: University-of-Seville.png
---

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)
library(conflicted)
library(ggplot2)
library(MKinfer)
library(formatR)
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::lag)
```

# Introduction
## Introduction
- Selection of stably expressed reference genes is crucial for qRT-PCR
- Proyects objective is finding the most stable genes and distinguishing the influence of cell lines and experimental conditions
- Dataset of CT values from 4 cell lines, 9 genes, and 3 experimental conditions
- Final objective: Graphical representation of variances in CT values for each gene based on cell lines and experimental conditions

# Editing the data
## First edit of the data
```{r message=FALSE}
#read all data
data <- read_csv2("GeneDataAllCellLinesNewFormat.csv")

#change the variable names to a usable format in R
names(data)<-str_replace_all(names(data), c(" " = "."))
print(data, n=6)
```
## First overview of the data
```{r message=FALSE}
#show the mean of each gene in each cell line and experimental condition to get a first
#overview
data %>%
  group_by(Cell.Line, Gene, Experimental.Condition) %>% 
  summarise(mean(CT.Value, na.rm = TRUE)) %>% 
  print(n=3)
```
## Retrieving data variances by cell line
```{r echo=FALSE}
#get the data that is affected by the different cell lines
varianceDataCellLine <- data %>%
  #group the different "sets" of data
  group_by(Cell.Line, Gene, Experimental.Condition) %>% 
  #filter the data to only get the normoxic data, which means it will not be affected
  #by the experimental condition
  filter(Experimental.Condition == "24h NX") %>% 
  #group the data by the different genes
  group_by(Gene) %>% 
  summarise(
    #calculate the standard deviation of the CT values for each gene
    standardDeviationByCellLine = sd(CT.Value, na.rm = TRUE), 
    #calculate the mean of the CT values for each gene
    meanByCellLine = mean(CT.Value, na.rm = TRUE), 
    #calculate the coefficient of variation for each gene
    coefficientOfVariationCellLine = standardDeviationByCellLine / meanByCellLine, 
    #calculate the minimum of the confidence interval for the coefficient of variation
    confIntMinCellLine = sdCI(CT.Value)$conf.int[1] / meanByCellLine, 
    #calculate the maximum of the confidence interval for the coefficient of variation
    confIntMaxCellLine = sdCI(CT.Value)$conf.int[2] / meanByCellLine 
  )
varianceDataCellLine
```
## Retrieving data variances by experimental condition
```{r echo=FALSE}
#get the data that is affected by the different Experimental Conditions
varianceDataExpCond <- data %>%
  #group the different "sets" of data
  group_by(Cell.Line, Gene) %>% 
  #first get the data in groups for each different cell line,
  #because we want to clear out the effect of the different cell lines
  mutate(
    #calculate the standard deviation of the CT values for each gene and each cell line
    standardDeviationByExpCond = sd(CT.Value, na.rm = TRUE), 
    #calculate the mean of the CT values for each gene and each cell line
    meanByExpCond = mean(CT.Value, na.rm = TRUE), 
    #calculate the coefficient of variation for each gene and each cell line
    coefficientOfVariationExpCond = standardDeviationByExpCond / meanByExpCond, 
    #calculate the minimum of the confidence interval for the coefficient of variation
    confIntMinExpCond = sdCI(CT.Value)$conf.int[1] / meanByExpCond, 
    #calculate the maximum of the confidence interval for the coefficient of variation
    confIntMaxExpCond = sdCI(CT.Value)$conf.int[2] / meanByExpCond 
  ) %>% 
  #group the data by the different genes to get the means of all the values
  #for each gene (like this we clear out the effect of the different cell lines)
  group_by(Gene) %>% 
  summarise(
    #calculate the mean of the coefficient of variation for each gene
    coefficientOfVariationExpCond = mean(coefficientOfVariationExpCond, na.rm = TRUE), 
    #calculate the minimum of the confidence interval for the coefficient of variation
    confIntMinExpCond = mean(confIntMinExpCond, na.rm = TRUE), 
    #calculate the maximum of the confidence interval for the coefficient of variation
    confIntMaxExpCond = mean(confIntMaxExpCond, na.rm = TRUE) 
  )
varianceDataExpCond
```

## Merge the data
```{r message=FALSE}
#group all import data into a tibble
finalData <- merge(varianceDataCellLine, varianceDataExpCond, by = "Gene")
finalData <- as_tibble(finalData)
print(finalData, n=4)
```
## Plotting the raw data

```{r}
ggplot(data = finalData) +
  aes(x = coefficientOfVariationExpCond, y = coefficientOfVariationCellLine, color = Gene) +
  geom_point()
```

## Adding error bars

```{r, echo=FALSE}
ggplot(data = finalData) +
  aes(x = coefficientOfVariationExpCond, y = coefficientOfVariationCellLine, color = Gene) +
  geom_point() +
  geom_errorbar(aes(xmin = confIntMinExpCond, xmax=confIntMaxExpCond), width=.01) +
  geom_errorbar(aes(ymin = confIntMinCellLine, ymax=confIntMaxCellLine), width=.01)
```


## Optimizing the plot (error bars are too dense for a graphical representation)

```{r, echo=FALSE}
ggplot(data = finalData) +
  aes(
    x = coefficientOfVariationExpCond * 100,
    y = coefficientOfVariationCellLine * 100,
    color = Gene
  ) +
  geom_point() +
  geom_text(label=finalData$Gene, hjust=0, vjust=0) +
  lims(x = c(0, 7.5), y = c(0, 7.5)) +
  #add a dashed line to show the 1:1 line.
  #Any gene that is under this line means that the
  #coefficient of variation of this gene is higher by
  #the experimental conditions than by the cell lines
  geom_abline(intercept = 0, slope = 1, color = "grey", linetype = "dashed") +
  labs(
    x = "The coefficient of variation by experimental conditions in %",
    y = "The coefficient of variation by cell lines in %",
    title = "Coefficient of variation of the CT-Values for the different analyzed genes"
  )
```